\documentclass[12pt]{article}                         % Make it an article
\usepackage{indentfirst}                       % All paragraphs indented
\usepackage{setspace}                           % For 1.5 spacing
\usepackage[hidelinks]{hyperref}      % Make table of contents clickable
\usepackage[margin=1in]{geometry} % 1 inch margin
\usepackage[english]{babel}                     % For "fancy" quotes
\usepackage[autostyle, english=american]{csquotes} % Same
\usepackage{apacite}                            % Use APA citations
\usepackage{etoolbox}                           % Misc
\usepackage{titlesec}
\usepackage{listofauthorships}
\MakeOuterQuote{"}                              % For fancy quotes
\usepackage{times}                              %For Tilde
\usepackage{textcomp}                           %For Tilde
\titlespacing*{\section}{0pt}{0pt}{0pt}
\titlespacing*{\subsection}{0pt}{0pt}{0pt}
\titlespacing*{\subsubsection}{0pt}{5pt}{-3pt}

\title{Eilat Transportation}
\author{Evan Goldstein, Valeria Kopper, Christopher Myers, Zachary Zlotnick}
\date{January 2018}

\begin{document}
\pagenumbering{gobble}
\maketitle
\newpage

\renewcommand\abstractname{Summary} % "Abstract" -> "Summary", the header lies

\pagenumbering{roman}
\tableofcontents
\newpage
\listofauthorships
\newpage
\pagenumbering{arabic}
\doublespacing

% Actual content begins here!
\section{Introduction}[All]

Eilat is a city in the southern tip of Israel with a growing population of ~50,000 and an area of 85 km$^2$. The city is a popular tourist destination, with over 3 million tourists per year from around the world. Tourism is expected to continue increasing in the future.

The city currently suffers from heavy congestion at peak hours, and recent regional transport changes are expected to worsen the situation, increasing travel times. This compounding congestion puts a strain on Eilat's transport network, calling for an increased effort to shorten commute times, expand Eilat's transportation services, and streamline the travel experience.

Understanding current constrains for transportation in Eilat will allow for planning what future transportation will look like in the city, and the implementation of certain smart city initiatives. This will optimize city functions as well as improve life quality for residents, tourists, and businesses.

Current transportation offerings include Egged intra-city buses in Eilat. A bus journey from Yotam Square (on the west side of the city) to the airport takes about 26 minutes with no traffic, and from Yotam Square to the Hotels Zone takes 35 minutes with no traffic. A bus journey from Mosh Beach to Big Eilat also takes around 35 minutes, but each of these trips could take less than 10 minutes with a better system. Other transportation options such as trams have not been previously considered, and could replace buses or work alongside them to provide more direct routes and decrease travel times within the city.

%Both these options could use dedicated tracks, and would not suffer from road congestion during peak travel times.

The goal of this project is to aid the city of Eilat in investigating possible transportation services that will minimize environmental impact, congestion, and travel times for residents, tourists, and businesses while improving ease of travel. To accomplish this, we will first develop a detailed understanding of Eilat's current transportation network, its advantages, and disadvantages. We will then create predictions for future congestion in the city based on projected growth and currently planned transportation changes. We will produce a roadmap to guide Eilat's transportation development through 2040 and present data on current and future transport. An optimized transportation network will improve resident quality of life and improve tourism and the economy in Eilat.
 
\newpage
\section{Background}

\subsection{Existing Transportation}[Chris M., Zach Z.]
Israel's route 90 is the main highway leading into Eilat, positioned on its northern edge and running parallel to the current Eilat Airport. The Ovda Airport is scheduled to shut down as the new Ramon Airport opens alongside route 90, and Ovda Airport is expected to cease handling civilian flights. This leaves Ramon airport as the only airport servicing Eilat, and therefore route 90 as the only road handling tourist traffic. Route 90 does not appear set up for this, with only a simple road junction for the airport and a series of traffic circles in the city itself. With already present congestion in Eilat near the route's entrance to the city and tourist traffic from the new airport, overall heavy congestion and increased travel times are expected.

Alternatives to conventional public methods of transportation already exist to bridge the gaps in commuting. "Bird" has successfully implemented an electric scooter system in over 100 cities worldwide --- solving public transport's "last mile" problem. Riders pick up a scooter from any "nest" in the city and pay an initial fee plus a fixed fee for each minute used \cite{MichaelRaz-Chaimovich2018BirdGlobes}. Bird riders usually don't exceed 3km per use, suggesting short-range use only, for example to or from bus stops. \cite{MichaelRaz-Chaimovich2018BirdGlobes}. Scooters in Tel Aviv are used heavily to avoid difficult road traffic. Another alternative is Mobike, which provides similar ease-of-access bike transportation, but at much lower cost, both in fees and maintenance. Finally, another option is CAR2GO, which allows users to rent a car for any time, but without requiring human intervention or paper signing. The cost of parking and insurance is part of the rental fee, and the experience is fairly seamless for users \cite{OrenDoriandMeiravMoran2018IsraeliHaaretz.com}. CAR2GO is used in 6 Israeli cities and could be used in Eilat, appealing to tourists with its ease of use.

\subsection{Airports near Eilat}[Evan G.]
In 2017, 1.4 million passengers passed through the J. Hozman airport in Eilat [iaa], and another 200 thousand passed through Ovda International Airport [iaa] for a total of 1.6 million passengers between the two airports. Ramon Airport will replace both the Eilat and Ovda airports, handling both domestic and international flights [citation needed]. Ramon will accommodate up to 2.25 million passengers a year by 2020, and 4.25 million passengers a year by 2030 [iaa]. This suggests a minimum of over 600 passengers per hour, assuming flights landing between 06:00 and 00:00 (an 18 hour schedule) [citation needed]. Transporting this many people to and from the city will require a transportation method that minimizes congestion, environmental impact, and travel time.

\subsection{Trains \& Trams}[Evan G., Chris M.]
Trains are a good option for minimizing congestion, as they run on dedicated tracks instead of roads [citation?]. A rail line between the airport and the city could be integrated with the existing line between Tel Aviv and Jerusalem. Magnetic Levitation, or Maglev, is the fastest type of train available, but also the most expensive, costing between 50 and 200 million dollars per mile [citation needed]. Light rail is the most common type of rail transit used for short-range, often intra-city, transport, but typically average between 20-30 mph[Lighrailnow]. Metrail, a hybrid solution, is a lower cost and environmentally friendly option, and can be installed for only 20 million dollars per mile[Metrail]. A compromise between cost, speed and environmental impact is a more traditional rail system with trains such as Bombardier's TALENT 3, which has a cost of less than 7.5 million dollars per train car and speeds of 160 kph[railway-technology]. The TALENT 3 uses traditional rails, which have an installation cost of between 1 and 2 million dollars per mile [acwr].

Another option for transportation within Eilat is a tram system. Trams are small rail-based vehicles that run along tracks in designated areas, or through roads for crossing, allowing the tram to act as a street vehicle. Trams combine the capacity and efficiency of a train with the accessibility and cost per unit of a bus. Trams can stop at any point and take on passengers without a station and are an efficient use of space on or off a road. Trams are typically all-electric, with power delivered by one or two overhead lines strung well above the tracks, making trams both energy efficient and zero-emission. If a train line to the new airport was established, the tram network could theoretically be connected to that too, making trams an easy connection both within and outside the city. Tram systems have been well-tested, with hundreds across the globe carrying billions of passengers per year, including one in Jerusalem.

Tram systems can be challenging to set up, so some intermediary steps may be advised. Trams require rails to run on, posing an up-front infrastructure cost beyond the cost of vehicles alone. High-voltage wires to deliver power are also needed but are likely far cheaper than embedding rails in roads, so an intermediary system of trolleybuses (electric buses powered by overhead lines) may be useful to ease the transition and ensure its viability before completely committing to it. 

\subsection{Smart Cities}[Valeria K.]

Through the use of innovative technologies that incorporate data analysis, cities have begun developing an interconnected web, increasingly through the Internet of Things (IoT), that allows for maximized city function efficiency \cite{MargaretRouseSmartCity}. These are known as smart cities, which can lead to economic growth and better life quality for citizens. Successful smart cities tend to have strong support from their government, especially in regards to open data platforms and smart sensors, as well as collaboration between the private and the public sectors \cite{BrianZanghi2017WhyExamples}. Oftentimes transparency, achieved by making collected data available to citizens, greatly increases confidence, which plays a role in the successful development of smart cities.

A combination of technologies, such as sensors throughout the city, can lead to improved traffic flow with fewer traffic jams --- a result of smart traffic management. Through monitoring and analyzing traffic flow, cities can better adjust and manage public transportation and traffic lights. Waze, a navigation app, has partnered with several cities in order to form a two-way exchange of information called the Connected Citizens Program \cite{Stern2016WazeMobility}. Other cities have implemented smart meters, which show users in an app where they can find free parking spaces; it also allows for digital payment. Furthermore, existing smart cities have found that smart public transit improves the efficiency and the satisfaction of users by fulfilling the real time demand. Through smart sensors there can be a significant energy conservation since street lights can be dimmed when there are no pedestrians in the area. Overall, the technology found in smart cities improves life quality.

\newpage
\section{Methods}

\subsection{Understanding Transportation}[Valeria K.]

In order to develop a plan for Eilat's future transportation we must first have a comprehensive understanding of Eilat's current transportation situation. Through observational research we will gather information and a better understanding of the city's layout and overall transportation trends. Understanding the interaction between residents and tourists will be crucial; furthermore, understanding how these two factors affect traffic flow on a seasonal scope will be important. Analyzing statistics about travel to Eilat can reveal important information that can then be developed into graphics to better display information.

\subsection{Eilat's Future}[Chris M.]

To develop plans and predictions we will use a combination of data gathered in the previous stage, information about Eilat publicity, and information about Eilat's expansion over time. On a basic data analysis level we can review data collected about traffic patterns, population, and city area over time to model these patterns and generate long-term predictions. We can investigate Eilat's tourism patterns with relation to its online presence and publicity by gathering data from sources like Google Trends. Google Trends provides data on searches for key words or topics over time, so we can gauge tourism interest at present and in the future. Increased tourism implies increased traffic, which we can use to help generate future traffic heatmaps with the help of other mapping and traffic tools like Google Maps or Waze. As part of our investigation we will also look for areas that might be ideal for new bus or tram stops and determine the viability of installing tram systems, plus options for small transport like bikes or scooters.

\subsection{Transportation Roadmap}[Evan G., Zach Z.]
With the data we will have collected, we will determine which transportation options best suit the city, assessing the shortcomings of the existing system and using growth predictions to assess future transportation needs. We will look at the expectations for traffic to and from Ramon Airport, and determine whether mass transit options between the airport and the city will be necessary to mitigate congestion within the city.

We will create a roadmap that will exhibit a timeline or adjusted city scheme, showing updated transportation systems from the Ramon Airport into the city, suggestions for new transport options or routes if applicable, and areas of high congestion. The roadmap will take into account the time of year and time of day in its analysis to highlight key congestion points, to help ensure that public transport is not over-stressed in peak hours or oversupplied during standard hours. This may involve creating suggestions for schedules or deployments if applicable.

% Bibliography -- automatically managed in APA style, on a new page
\newpage
\bibliographystyle{apacite}
\bibliography{refs}

\end{document}
